The Efficient Handling of Guards \\in the Design of RPython's Tracing JIT

### 2012-08-17 talk/dls2012/paper.tex
too much capitalization

\begin{figure}
\begin{lstlisting}[mathescape,numbers = right,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
$L_0$($p_{0}$, $p_{1}$):
# inside f: y = y.add(step)
guard_class($p_{1}$, BoxedInteger)
    # inside BoxedInteger.add
    $i_{2}$ = get($p_{1}$, intval)
    guard_class($p_{0}$, BoxedInteger)
        # inside BoxedInteger.add__int
        $i_{3}$ = get($p_{0}$, intval)
        $i_{4}$ = $i_{2} + i_{3}$
        $p_{5}$ = new(BoxedInteger)
            # inside BoxedInteger.__init__
            set($p_{5}$, intval, $i_{4}$)
jump($L_0$, $p_{0}$, $p_{5}$)
\end{lstlisting}
\caption{An Unoptimized Trace of the Example Interpreter}
\label{fig:unopt-trace}
\end{figure}

@@

$L_1$($p_{0}$, $i_{3}$, $i_{4}$):
    $i_{8}$ = $i_{4}+i_{3}$
jump($L_1$, $p_{0}$, $i_{3}$, $i_8$)
\end{lstlisting}
\caption{The fully optimized loop of the Example Interpreter}
\label{fig:opt-trace}
\end{figure}

@@

The speedups that LuaJIT gains from the loop optimization pass are similar to
those PyPy gains. In general, LuaJIT is even closer to C performance, sometimes
even surpassing it. LuaJIT is generating machine code of higher quality because
it has more optimizations\footnote{See
\texttt{http://wiki.luajit.org/Optimizations}} and is a produces much better
machine code than PyPy.

>>>>

\begin{figure}
\begin{lstlisting}[mathescape,numbers = right,basicstyle=\setstretch{1.05}\ttfamily\scriptsize]
$L_0$($p_{0}$, $p_{1}$):
# inside f: y = y.add(step)
guard_class($p_{1}$, BoxedInteger)
    # inside BoxedInteger.add
    $i_{2}$ = get($p_{1}$, intval)
    guard_class($p_{0}$, BoxedInteger)
        # inside BoxedInteger.add__int
        $i_{3}$ = get($p_{0}$, intval)
        $i_{4}$ = $i_{2} + i_{3}$
        $p_{5}$ = new(BoxedInteger)
            # inside BoxedInteger.__init__
            set($p_{5}$, intval, $i_{4}$)
jump($L_0$, $p_{0}$, $p_{5}$)
\end{lstlisting}
\caption{An Unoptimized Trace of the example interpreter}
\label{fig:unopt-trace}
\end{figure}

@@

$L_1$($p_{0}$, $i_{3}$, $i_{4}$):
    $i_{8}$ = $i_{4}+i_{3}$
jump($L_1$, $p_{0}$, $i_{3}$, $i_8$)
\end{lstlisting}
\caption{The fully optimized loop of the example interpreter}
\label{fig:opt-trace}
\end{figure}

@@

The large input parameters of the SciMark benchmarks are chosen in such a way
to make the problem not fit into the CPU cache. This explains why PyPy is doing
relatively better on them. The cache miss penalties are large relative to the
time needed to perform the actual computations, which hides problems of the
less efficient code generated by PyPy.

The speedups that LuaJIT gains from the loop optimization pass are similar to
those PyPy gains. In general, LuaJIT is even closer to C performance, sometimes
even surpassing it. LuaJIT is generating machine code of higher quality because
it has more optimizations\footnote{See
\texttt{http://wiki.luajit.org/Optimizations}} and produces much better
machine code than PyPy.

### 2012-08-17 talk/vmil2012/paper.tex
too much capitalization

RPython can mean one of two things:
\begin{itemize}
 \item the language itself
 \item the translation toolchain used to transform RPython programs to executable units
\end{itemize}
The RPython language, is a subset of Python that provides a
statically typed object-oriented high level language. The subset of Python available in RPython is chosen in a way type inference is possible\cite{ancona_rpython:_2007}.
The language provides
several features such as automatic memory management
and just-in-time compilation. When writing an interpreter using RPython the
programmer only has to write the interpreter for the language she is
implementing.  The second RPython component, the translation toolchain, is used
to transform the interpreter into a C program.\footnote{
    RPython can also be used to translate programs to CLR and Java
    bytecode~\cite{ancona_rpython:_2007}, but this feature is somewhat
    experimental.
}
During the transformation process
different low level aspects suited for the target environment are automatically
added to the program such as a garbage collector and a tracing JIT compiler.
The process of inserting a tracing JIT is not fully automatic but is guided by
hints from the interpreter author.

@@

Mike Pall, the author of LuaJIT describes in a post to the lua-users mailing
list different technologies and techniques used in the implementation of
LuaJIT~\cite{Pall:2009}. Pall explains that guards in LuaJIT use a datastucture
called snapshots, similar to RPython's resume data, to store the information
about how to rebuild the state from a guard failure using the information in the
snapshot and the machine execution state. According to Pall~\cite{Pall:2009}
snapshots for guards in LuaJIT are associated with a large memory footprint.
The solution used there is to store sparse snapshots, avoiding the creation
of snapshots for every guard to reduce memory pressure. Snapshots are only
created for guards after updates to the global state, after control flow points
from the original program and for guards that are likely to fail. As an outlook
Pall mentions plans to switch to compressed snapshots to further reduce
redundancy.\footnote{This optimization is now implemented in LuaJIT, at the time of writing it has not been fully documented in the LuaJIT Wiki (\url{http://wiki.luajit.org/Optimizations\#1-D-Snapshot-Compression}).}
The approach of not creating snapshots at all for every guard is
orthogonal to the resume data compression presented in this paper and could be
reused within RPython to improve the memory usage further.

>>>>

RPython can mean one of two things:
\begin{itemize}
 \item the language itself
 \item the translation toolchain used to transform RPython programs to executable units
\end{itemize}
The RPython language is a
statically typed object-oriented high-level subset of Python. The subset is chosen in such a way to make type inference possible\cite{ancona_rpython:_2007}.
The language provides
several features such as automatic memory management
and just-in-time compilation. When writing an interpreter using RPython the
programmer only has to write the interpreter for the language she is
implementing.  The second RPython component, the translation toolchain, is used
to transform the interpreter into a C program.\footnote{
    RPython can also be used to translate programs to CLR and Java
    bytecode~\cite{ancona_rpython:_2007}, but this feature is somewhat
    experimental.
}
During the transformation process
different low level aspects suited for the target environment are automatically
added to the program such as a garbage collector and a tracing JIT compiler.
The process of inserting a tracing JIT is not fully automatic but is guided by
hints from the interpreter author.

@@

Mike Pall, the author of LuaJIT describes in a post to the lua-users mailing
list different technologies and techniques used in the implementation of
LuaJIT~\cite{Pall:2009}. Pall explains that guards in LuaJIT use a datastucture
called snapshots, similar to RPython's resume data, to store the information
about how to rebuild the state from a guard failure using the information in the
snapshot and the machine execution state. According to Pall~\cite{Pall:2009}
snapshots for guards in LuaJIT are associated with a large memory footprint.
The solution used there is to store sparse snapshots, avoiding the creation
of snapshots for every guard to reduce memory pressure. Snapshots are only
created for guards after updates to the global state, after control flow points
from the original program and for guards that are likely to fail. As an outlook
Pall mentions plans to switch to compressed snapshots to further reduce
redundancy.\footnote{This optimization is now implemented in LuaJIT, at the time of writing it has not been fully documented in the LuaJIT Wiki: \url{http://wiki.luajit.org/Optimizations\#1-D-Snapshot-Compression}}
The approach of not creating snapshots at all for every guard is
orthogonal to the resume data compression presented in this paper and could be
reused within RPython to improve the memory usage further.
